---
title: "Lecture 15: Spatial Prediction and Model Selection"
output:
  revealjs::revealjs_presentation:
    theme: white
    center: true
    transition: none
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
library(ggplot2)
library(leaflet)
```

# Class Intro

## Intro Questions 
- Spatial Generalized Linear Models: Explain how the spatial structure is incorporated into these models.

- For Today:
    - Spatial Prediction and Model Selection

# Kriging Predictions

## Multivariate Normal Theory
- For consider partioning a multivariate normal distribution into two parts
$$\begin{pmatrix}
\boldsymbol{Y_1}\\
\boldsymbol{Y_2}
\end{pmatrix} = 
N \left( \begin{pmatrix}
\boldsymbol{\mu_1}\\
\boldsymbol{\mu_2}
\end{pmatrix}, \begin{pmatrix}
\Omega_{11} \;\;\Omega_{12}\\
\Omega_{21} \;\;\Omega_{22}\end{pmatrix}\right),$$
where $\Omega_{12} = \Omega_{21}^T$

## Conditional Multivariate Normal Theory
- The conditional distribution, $p(\boldsymbol{Y_1}|
\boldsymbol{Y_2})$ is normal with:

- $E[\boldsymbol{Y_1}|
\boldsymbol{Y_2}] = \boldsymbol{\mu_1} + \Omega_{12} \Omega_{22}^{-1} (\boldsymbol{Y_2} - \mu_2)$

- $Var[\boldsymbol{Y_1}|
\boldsymbol{Y_2}] = \Omega_{11} - \Omega_{12} \Omega_{22}^{-1} \Omega_{21}$

- Typically, $\boldsymbol{\mu_1}$ requires an estimate of $\boldsymbol{\beta}$ and likewise $\Omega$ is a product of $\sigma^2$, $\phi$, and $\tau^2$.

- Strictly speaking we can write this distribution as $p(\boldsymbol{Y_1}|
\boldsymbol{Y_2}, \boldsymbol{\beta}, \sigma^2, \phi, \tau^2)$

# Cross Validation Framework

## Cross Validation Motivation

```{r}
load('CO_Air.Rdata')
pal <- colorFactor(c("green", "red"), domain = c(FALSE, TRUE))
leaflet(CO) %>% addTiles() %>% 
  addCircleMarkers(lng = ~Longitude, lat = ~Latitude,
    color = ~pal(Exceedance)) %>%  addCircleMarkers(lng = ~Longitude[1], lat = ~Latitude[1], radius = 15)
```

## Prediction
- We will use cross-validation or a test/training approach to compare predictive models.
- Consider three data structures: continuous, count, and binary; how should we evaluate predictions in these situations?
